Constructing a universe for the setoid model

### 2021-01-21 paper.tex
Clarification on use of weird identity type.
fixes suggested by reviewer 2, saving some space when rewriting the intro of simple eliminators

Finally, we will assume an identity type in the style of Martin-L\"of's
inductive identity type. The main difference is that our identity type is a
$\mProp$-valued relation. We have a transport combinator
\textsf{transp} from which \textsf{J} is derivable.
\vspace{-0.2em}
\begin{align*}
  \textsf{Id} &: \{A : \mType \} \to A \to A \to \mProp \\
  \textsf{refl} &: \{A : \mType \} (a : A) \to \textsf{Id}\ a\ a \\
  \textsf{transp} &: \{A : \mType \} (C : A \to \mType) \{a_0\ a_1 : A\} \to \textsf{Id}\ a_0\ a_1 \to C\ a_0 \to C\ a_1
\end{align*}
with $\textsf{transp}\ C\ \{x\}\ \{x\}\ e\ u \equiv u$.
The \textsf{transp} combinator provides a strong elimination principle allowing
to eliminate a strict proposition (the identity type) into arbitrary types.
%
We only use this identity type with such a strong transport in Section~\ref{inductive-universe}.
%
% Note that this identity type must necessarily be assumed as
% primitive; in fact, even though we could have defined \textsf{Id}
% from $W_\mProp$, \textsf{transp} is not definable from $W_\mProp$
% elimination.

@@

We recall the rules of SeTT that extend regular MLTT below, but with a
variation: whereas the equality types in \cite{mpc19} are stated as elements of
SeTT's internal universe of propositions, here we state the context equalities
as elements of the external, metatheoretic universe $\mProp$. This makes it easier
to define models of SeTT as contexts do not need to be partially fibrant. Equality on
types is defined as before in \cite{mpc19}.

@@

Yet another choice is whether to have two separate sorts, one for propositions
and one for sets (with propositions convertible to sets) or a single sort of
types (sets), with propositions given by elements of a universe of proposition,
which is a (large) type. We have chosen to present the second option to fit with
the standard notion of (unisorted) CwF. However, this has downsides: to even
talk about propositions, we need to have a notion of large types. The first
option is more symmetric: we can have parallel hierarchies for propositions and
sets.

@@

For any inductive-inductive type $A : \mType, B: A \to \mType$, the most general
form of (dependent) eliminator is given by motives $F : A \to \mType$ and $G :
(a : A) \to (b : B a) \to F\ a \to \mType$, and the functions $\textsf{elim}_A :
(a : A) \to F\ a, \textsf{elim}_B : (a : A) \to (b : B\ a) \to
G\ a\ b\ (\textsf{elim}_A\ a)$. A simpler notion of eliminator does not require
the second component of the motive to mention the first. In particular, we have
$F : A \to \mType$, $G : (a : A) \to B\ a \to \mType$, and eliminators
$\textsf{elim}_A : (a : A) \to F\ a, \textsf{elim}_B : (a : A) (b : B\ a) \to
G\ a\ b$. Note that this elimination principle---that we call \emph{simple
  elimination} after \cite{ii}---is no longer recursive-recursive, since
$\textsf{elim}_A$ is not mentioned in the signature of $\textsf{elim}_B$.

Let us go back to our \textsf{Con}/\textsf{Ty} example. Given any dependent
algebra of $\textsf{Con}$ and $\textsf{Ty}$ with motives $C : \textsf{Con} \to
\mType$ and $T : (\GG : \textsf{Con}) \to \textsf{Ty}\ \GG \to C\ \GG \to
\mType$, the general eliminators have the following signatures:
\begin{alignat*}{10}
  & \textsf{elim}_{\textsf{Con}} && : (\GG : \textsf{Con}) \to C\ \GG \hspace{5em} \\
  & \textsf{elim}_{\textsf{Ty}} && : \{\GG : \textsf{Con}\} (A : \textsf{Ty}\ \GG) \to T\ \GG\ A\ (\textsf{elim}_{\textsf{Con}}\ \GG)
\end{alignat*}

These can be derived from our encoding of $\textsf{Con}$ and $\textsf{Ty}$ via
untyped codes and well-typing predicates. The way to do it is to first define
the graph of the eliminators in the form of inductively-generated relations:
\begin{alignat*}{10}
  & \textsf{data}\ \textsf{R-Con} && :
    (\GG : \textsf{Con}) \to C\ \GG \to \mType \\
  & \textsf{data}\ \textsf{R-Ty} && : \{\GG : \textsf{Con}\} (A : \textsf{Ty}\ \GG)
  (\gamma : C\ \GG)
  \to T\ \GG\ A\ \gamma
  \to \mType
\end{alignat*}

>>>>

Finally, we will assume an identity type in the style of Martin-L\"of's
inductive identity type. The main difference is that our identity type is a
$\mProp$-valued relation. We have a transport combinator
\textsf{transp} from which \textsf{J} is derivable.
\vspace{-0.2em}
\begin{align*}
  \textsf{Id} &: \{A : \mType \} \to A \to A \to \mProp \\
  \textsf{refl} &: \{A : \mType \} (a : A) \to \textsf{Id}\ a\ a \\
  \textsf{transp} &: \{A : \mType \} (C : A \to \mType) \{a_0\ a_1 : A\} \to \textsf{Id}\ a_0\ a_1 \to C\ a_0 \to C\ a_1
\end{align*}
with $\textsf{transp}\ C\ \{x\}\ \{x\}\ e\ u \equiv u$.
The \textsf{transp} combinator provides a strong elimination principle allowing
to eliminate a strict proposition (the identity type) into arbitrary types.
%
We only use this identity type in Section~\ref{inductive-universe}.
For the rest of our constructions, the traditional Martin-L\"of's identity type suffices.

%
% Note that this identity type must necessarily be assumed as
% primitive; in fact, even though we could have defined \textsf{Id}
% from $W_\mProp$, \textsf{transp} is not definable from $W_\mProp$
% elimination.

@@

We recall the rules of SeTT that extend regular MLTT below, but with a
variation: whereas the equality types in \cite{mpc19} are stated as elements of
SeTT's internal universe of propositions, here we state the context equalities
as elements of the external, metatheoretic universe $\mProp$. This generalises the
notion of model of SeTT thus making it easier to construct models. Equality on
types is defined as before in \cite{mpc19}.

@@

Yet another choice is whether to have two separate sorts, one for propositions
and one for sets (with propositions convertible to sets) or a single sort of
types (sets), with propositions given by elements of a universe of propositions,
which is a (large) type. We have chosen to present the second option to fit with
the standard notion of (unisorted) CwF. However, this has downsides: to even
talk about propositions, we need to have a notion of large types. The first
option is more symmetric: we can have parallel hierarchies for propositions and
sets.

@@

Following \cite{ii}, we distinguish two versions of the eliminator:
the simple and the general one. Note that this is orthogonal to the
distinction between nondependent and dependent eliminators, from which
we only consider the latter. The motives for the simple eliminator are
$C':\textsf{Con}\to\mType$,
$T':(\GG:\textsf{Con})(A:\textsf{Ty}\,\GG)\to \mType$ and the
eliminators themselves have the following signatures:
\begin{alignat*}{10}
  & \textsf{elim}_{\textsf{Con}}' && : (\GG : \textsf{Con}) \to C'\ \GG \hspace{3.3em} \textsf{elim}_{\textsf{Ty}}' && : \forall\{\GG\} (A : \textsf{Ty}\ \GG) \to T'\ \GG\ A \hphantom{(\textsf{elim}_{\textsf{Con}}\ \GG)}
\end{alignat*}
In the case of the general eliminator, the motive for $\textsf{Ty}$
depends on the motive for $\textsf{Con}$, making the two eliminators
\emph{recursive-recursive} functions. For motives $C :
\textsf{Con} \to \mType$ and $T : (\GG : \textsf{Con}) \to
\textsf{Ty}\ \GG \to C\ \GG \to \mType$ the signatures are:
\begin{alignat*}{10}
  & \textsf{elim}_{\textsf{Con}} && : (\GG : \textsf{Con}) \to C\ \GG \hphantom{'}\hspace{3.3em} \textsf{elim}_{\textsf{Ty}} && : \forall\{\GG\} (A : \textsf{Ty}\ \GG) \to T\ \GG\ A\ (\textsf{elim}_{\textsf{Con}}\ \GG)
\end{alignat*}

The general eliminators can be derived from our encoding of $\textsf{Con}$ and $\textsf{Ty}$ via
untyped codes and well-typing predicates. The way to do it is to first define
the graph of the eliminators in the form of inductively-generated relations:
\begin{alignat*}{10}
  & \textsf{data}\ \textsf{R-Con} && :
    (\GG : \textsf{Con}) \to C\ \GG \to \mType \\
  & \textsf{data}\ \textsf{R-Ty} && : \{\GG : \textsf{Con}\} (A : \textsf{Ty}\ \GG)
  (\gamma : C\ \GG)
  \to T\ \GG\ A\ \gamma
  \to \mType
\end{alignat*}
